Sketch Recogniton using Particle Swarm Algorithm

### 2009-04-12 trunk/draft/System.tex
now is 11 april
before i change the results

\chapter{The Proposed System}
\label{sec:proposedSystem}
\section{Aims of the System}
\label{sec:AimsOfTheSystem}
The previous chapter showed the current trends in sketch recognition. This thesis introduces a new approach for the sketch recognition problem. Particle Swarm Optimization (PSO) is used to correctly segment strokes into curves and lines. Particle Swarm Optimization (PSO) is used to correctly segment strokes into curves and lines. The current implemented systems either segment strokes based on simple curvature\cite{meanshift10,earlySketchbased4} or speed information\cite{earlySketchbased4}. %the performance of most algorithms is based on user style.

@@

\section{System Overview}
\label{sec:AnOverviewOfTheSystem}
   This research solves sketch segmentation problem using particle swarm algorithm PSO.  Due to sloppiness of users and hardware glitches the captured points must be processed to remove noise and redundancy before the system proceeds to compute the speed and curvature information. After that a set of possible corner or critical points is computed to guide the segmentation procedure.  A set of features is extracted from the segmented strokes to be used by the classifier.  %The stroke is the path of points from the instant the pen is down till it is up.
The block diagram in Figure \ref{fig:Blockdiagram} shows the main system blocks. The next section describes each block in details.
 %Due to sloppiness of users and hardware glitches the system processes the data before further
\begin{figure}[]
	\centering

\begin{center}
	\includegraphics[scale=0.75]{images/AllBlockDiagram.eps}
	\caption[The System Block Diagram]{The block diagram of the system}
	\label{fig:Blockdiagram}
\end{center}
\end{figure}

@@

The \textbf{preprocessing} step is  is responsible for capturing the input data and removing the noise from it. As shown in Figure \ref{fig:Blockdiagram} preprocessing consists of computing the time differences, direction, speed and curvature of every point in the stroke. After computing this information the local minimum and local maximum of each curve is extracted. This process estimates a set of a set of Possible Dominant Points $P_{pd}$ that guide the segmentation process. Later, other geometrical and statistical computations are evaluated from the stroke. %The system then proceeds to estimate to help in the segmentation process.   \\

@@

After each segment type is determined the the particle segmentation error is computed by Equation \ref{eq:errorSwarm2} 
\begin{equation}
E=\sum\nolimits_{i = 0}^M e(D_i) 
\label{eq:errorSwarm2}
\end{equation}
where $M$ is the number of segments in the solution, $D_i$ is the  approximation error of each segment as $min(d_c,d_s)$ as computed in Equation\ref{eq:circleE}  and Equation \ref{eq:ds} \cite{CruveDivisionSwarm}. The fitness is computed by the equation \ref{eq:fitnessSwarm2} 
\begin{equation}
\max fitness(P_i ) = \frac{1}{{E \times M^k }}
\label{eq:fitnessSwarm2}
\end{equation} where $E$ is the error and $M$ is number of segments and $k$ is a parameter tweaked to get minimum number of segments. $k$ is selected to be 0.5\cite{CruveDivisionSwarm}.

\subsubsection{Solution Refine procedures } 
%\cite{PolygonApproximationPSO} used a merge and divide algorithm after each loop of the swarm system to refine the solution but we used another enhancement method.
 After each loop in the swarm algorithm, each particle loops on the set of selected dominant points to enhance the solution. Each dominant point is checked to find if it was labeled before as a possible dominant (computed as in section \ref{sec:Preprocessing}), if not the point is moved to the nearest labeled point.
   After each loop of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedure. For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{sec:Preprocessing}). If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point. This ensures that all of the points generated by the DPSO are possible dominate points $P_{pd}$. After that the particles are tested to make sure that the distance between every two successive dominate point is larger than the constant $min_D$. If two points are nearer than $min_D$\footnote{Constant selected to be 10 pixels} then one of the points is removed. 
\subsection{Choosing the best fit}
The two PSO algorithms produce two segmentation solutions. After that the system evaluates each solution to finally select the best segmentation. This evaluation is based on the area of both the segmentation and the input stroke. Equation \ref{eq:bestFit1} and \ref{eq:bestFit2} are used to evaluate each segmentation.

@@

 \section{Hybird Segmentation Algorithm}

\label{sec:BenchMarckAlgorithm}
In this research, we developed another algorithm as reference to the results obtained by the DPSO algorithms. The algorithm was first implemented by  Sezgin et al. in \cite{earlyprocess}. The algorithm is divided into the following steps: 
\begin{enumerate}
	\item Compute the curvature and the speed for each point in the stroke. 
	\item   A percentage of maximum value of the speed and curvature is used as threshold.
		\item Mark the values higher than the threshold in a new arrays $L_c$ as the list of corner points generated from the curvature curve. The list $L_s$ contains the points generated from the speed curve.
	\item Compute a confidence on each list based on the type of curve. Then sort the lists $L_s$and $L_c$ by the confidence value. 
	\item Find the intersection points in both list and generate the a Sample solution $h_{curr}$. 
	\item Use $h_{curr}$ as initial solution. Loop doing the following:

@@

The system uses a composite set of features include Rubine feature set \cite{gestureexample12}, Zernike moments \cite{HeloiseBeautification}, ink density \cite{GeometryAndDomain102} and some structural information like number of perpendiculars lines , number of parallel lines and types of primitives in each symbol. The table \ref{tab:FeatureTable} gives details on all the features used in the system.

\begin{description}
 \item [Structural and geometrical Features(FS1)] Features defines the structure of the geometrical symbol.  
\begin{description}
	 \item[No. Of segments] Number of segments in the symbol. 
		\item[No. Of primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
		\item [No. of curves] Number of curves or ellipses in the symbol. 
		\item [No. of lines] Number of lines in the symbol. 
		\item [No. of perpendicular lines] Number of perpendicular lines. 
		\item [No. of parallel lines] Number of parallel lines. 
		\item [No. of intersections] Number of intersection between lines and curves. 
		\item [Size Ratio] Ratio between width to height of the symbol.
		%\item[No. of holes] Number of holes in the symbols.

@@

\item[Statistical Features (FS3)]   
\begin{description}
	\item[Zernike moments]  Zernike moments of order \cite{HeloiseBeautification}. 
\end{description}
	\item[Composite Features (FS4)] Features that are composites of different geometrical and statistical symbol. 
	\begin{description}

>>>>

\chapter{The Proposed System}
\label{sec:proposedSystem}
\section{Aims of the System}
\label{sec:AimsOfTheSystem}
%The previous chapter showed the current trends in sketch recognition. 
This thesis introduces a new approach for the sketch recognition problem. Particle Swarm Optimization (PSO) is used to correctly segment strokes into curves and lines. The current implemented systems either segment strokes based on simple curvature\cite{meanshift10,earlySketchbased4} or speed information\cite{earlySketchbased4}. %the performance of most algorithms is based on user style.

@@

\section{System Overview}
\label{sec:AnOverviewOfTheSystem}
   This research solves sketch segmentation problem using particle swarm algorithm (PSO).  Due to sloppiness of users and hardware glitches the captured points must be processed to remove noise and redundancy before the system proceeds to compute the speed and curvature information. After that a set of possible corner or critical points is computed to guide the segmentation procedure.  A set of features is extracted from the segmented strokes to be used by the classifier.  %The stroke is the path of points from the instant the pen is down till it is up.
The block diagram in Figure \ref{fig:Blockdiagram} shows the main system blocks. The next section describes each block in details.
 %Due to sloppiness of users and hardware glitches the system processes the data before further
\begin{figure}[]
	\centering

\begin{center}
	\includegraphics[scale=0.75]{images/AllBlockDiagram.eps}
	\caption[The System Block Diagram]{Block diagram of the system}
	\label{fig:Blockdiagram}
\end{center}
\end{figure}

@@

The \textbf{preprocessing} step is  responsible for capturing the input data and removing the noise from it. As shown in Figure \ref{fig:Blockdiagram} preprocessing consists of computing the time differences, direction, speed and curvature of every point in the stroke. After computing this information the local minimum and local maximum of each curve is extracted. This process estimates a set of a set of Possible Dominant Points $P_{pd}$ that guide the segmentation process. Later, other geometrical and statistical computations are evaluated from the stroke. %The system then proceeds to estimate to help in the segmentation process.   \\

@@

After each segment type is determined the particle segmentation error is computed by Equation \ref{eq:errorSwarm2} 
\begin{equation}
E=\sum\nolimits_{i = 0}^M e(D_i) 
\label{eq:errorSwarm2}
\end{equation}
where $M$ is the number of segments in the solution, $D_i$ is the  approximation error of each segment as $min(d_c,d_s)$ as computed in Equation\ref{eq:circleE}  and Equation \ref{eq:ds} \cite{CruveDivisionSwarm}. The fitness is computed by the equation \ref{eq:fitnessSwarm2} 
\begin{equation}
\max fitness(P_i ) = \frac{1}{{E \times M^k }}
\label{eq:fitnessSwarm2}
\end{equation} where $E$ is the error and $M$ is number of segments and $k$ is a parameter tweaked to get minimum number of segments. $k$ is selected to be 0.5\cite{CruveDivisionSwarm}.

\subsubsection{Solution Refine procedures } 
%\cite{PolygonApproximationPSO} used a merge and divide algorithm after each loop of the swarm system to refine the solution but we used another enhancement method.
 After each loop in the swarm algorithm, each particle loops on the set of selected dominant points to enhance the solution. Each dominant point is checked to find if it was labeled before as a possible dominant (computed as in section \ref{sec:Preprocessing}), if not the point is moved to the nearest labeled point.
   After each loop of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedure. For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{sec:Preprocessing}). If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point. This ensures that all of the points generated by the DPSO are possible dominate points $P_{pd}$. After that the particles are tested to make sure that the distance between every two successive dominate point is larger than the constant $min_D$. If two points are nearer than $min_D$\footnote{Constant selected to be 10 pixels} then one of the points is removed.

\subsection{Choosing the best fit}
The two PSO algorithms produce two segmentation solutions. After that the system evaluates each solution to finally select the best segmentation. This evaluation is based on the area of both the segmentation and the input stroke. Equation \ref{eq:bestFit1} and \ref{eq:bestFit2} are used to evaluate each segmentation.

@@

\section{Hybird Segmentation Algorithm}
\label{sec:BenchMarckAlgorithm}

In this research, we developed another algorithm as reference to the results obtained by the DPSO algorithms. The algorithm was first implemented by  Sezgin et al. in \cite{earlyprocess}. The algorithm is divided into the following steps: 
\begin{enumerate}
	\item Compute the curvature and the speed for each point in the stroke. 
	\item   A percentage of maximum value of the speed and curvature is used as threshold.
		\item Mark the values higher than the threshold in a new arrays $L_c$ as the list of corner points generated from the curvature curve. The list $L_s$ contains the points generated from the speed curve.
	\item Compute a confidence on each list based on the type of curve. Then sort the lists $L_s$and $L_c$ by the confidence value. 
	\item Find the intersection points in both list and generate the a Sample solution $h_{curr}$. 
	\item Use $h_{curr}$ as initial solution. Loop on the following:

@@

The system uses a composite set of features include Rubine feature set \cite{gestureexample12}, Zernike moments \cite{HeloiseBeautification}, ink density \cite{GeometryAndDomain102} and some structural information like number of perpendiculars lines , number of parallel lines and types of primitives in each symbol. The following list gives details on all the features used in the system.

\begin{description}
 \item [Structural and geometrical Features(FS1)] Features defines the structure of the geometrical symbol.  
\begin{description}
	 \item[Segments] Number of segments in the symbol.
	 \item [Strokes] Number of strokes or partical strokes that created the symbols.  
		\item[Primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
		\item [Curves] Number of curves or ellipses in the symbol. 
		\item [Lines] Number of lines in the symbol. 
		\item [Perpendicular lines] Number of perpendicular lines. 
		\item [Parallel lines] Number of parallel lines. 
		\item [Intersections] Number of intersection between lines and curves. 
		\item [T intersections] Number of T intersections. 
		\item [L intersections] Number of L intersections. 
		\item [X intersections] Number of X intersections.
		\item [Size Ratio] Ratio between width to height of the symbol.

		%\item[No. of holes] Number of holes in the symbols.

@@

\item[Statistical Features (FS3)]   
\begin{description}
	\item[Zernike moments] Zernike moments of order n\footnote{ We tried different values of n. Results show the analysis of each order.} \cite{HeloiseBeautification}. 
\end{description}
	\item[Composite Features (FS4)] Features that are composites of different geometrical and statistical symbol. 
	\begin{description}

### 2009-04-12 trunk/draft/chapterSurvey.tex
before i change the results

 Avoiding graph matching problems, research tried to use languages that discribe shapes instead of graphs. Visual describtive languages represents the shape into tokens of geometric primitives or relations between smaller  geometric primitives. Figures \ref{fig:arrow1} and \ref{fig:arrow2} show a regular arrow with 2 shafts and head and its definition in Ladder language \cite{Ladder30}. Shape recognition is then a question of parsing the sketch document into a recognizer that parse strokes according to the constructed grammar.

 Several visual grammer were introduced \cite{statisticalparsing26,Ladder30,GenericHMM28}. \cite{Ladder30,GenericHMM28}  Used visual description language which will generate the information of the structure and constrain of each object classified. The system  automatically generates java code for the recognizer for each object. The generated code cycles through the strokes drawn to check if any of the recognizers generated had interpreted any symbols partially or completely.  These systems are less computaitonal expensive than graph matching approach but they also operate based on the output of the segmentation process. Any error in the segmentation process lead to huge error while parsing the whole sketch. Moreover, symbols cannot share strokes between them which infleunce and restrict how the user draw symbols. 
 % complexity of the graph matching problem 
\begin{figure}
	\centering
		\subfigure[An Arrow]
			{\label{fig:arrow1}\includegraphics[scale=0.75]{../../neededfiles/Figures/arrowladderdef1.eps}}
			\hfill
		\subfigure[Arrow grammar definition]
		 {	\label{fig:arrow2}\includegraphics[scale=0.88]{../../neededfiles/Figures/arrowladderdef2.eps}}
	\caption[Arrow definition]
	\label{fig:arrowladderdef1}
\end{figure}

@@

%Zernike moments were used by Hse et al. in \cite{}. 
Hse et al. \cite{HeloiseBeautification,zernike61} proposed using the magnitudes of Zernike moments as a shape property. Zernike moemnts are orthognals moments which define the distribution of point in the input space. The higher the order of the moment the finer the detail level the point distribution represents. Hse et al. \cite{HeloiseBeautification,zernike61} used a SVM classifier to train the features vector calculated from the moments. Early evluation of Zernike moemnts showed that they are invariant to both reflection and rotation of the input shape \cite{ZerMomentOrthogonal}. In our system we used Zernike moemnts as part of our feature vector. A detailed comparison between Zernike moments and other features are demonstrated in Section \ref{sec:featuresComparisions}.

@@

In all of these system only \cite{SketchRead2007} using stroke based recognitiion have been implemented to the fully creat a system that free draw sketch and understand them. This is because the emphasis on the interactive side of the sketch sytem which result in large difficultes in processing free hand sketches. Methods based on gesture and global stroke properties produce systems that are dependent on assumptions about how strokes can be groubed. These systems assume that users will finish drawing a shape before moving to the next one \cite{6multistrokeEvaluation,Cali63,geometrydomain49}. This let them to temporally group strokes and then recognize them based on their strokes or apperance properties. \cite{AlvaradoDigital} argued that this assumption dose not hold in natural sketchs.

@@

\section{Particle Swram Optimization Review}
\label{sec:ParticleSwramOptimizationReview}

%review on the pso alogrithm alone.................

>>>>

 Avoiding graph matching problems, research tried to use languages that discribe shapes instead of graphs. Visual describtive languages represents the shape into tokens of geometric primitives or relations between smaller  geometric primitives. Figures \ref{fig:arrow1} and \ref{fig:arr2} show a regular arrow with 2 shafts and head and its definition in Ladder language \cite{Ladder30}. Shape recognition is then a question of parsing the sketch document into a recognizer that parse strokes according to the constructed grammar.

 Several visual grammer were introduced \cite{statisticalparsing26,Ladder30,GenericHMM28}. \cite{Ladder30,GenericHMM28}  Used visual description language which will generate the information of the structure and constrain of each object classified. The system  automatically generates java code for the recognizer for each object. The generated code cycles through the strokes drawn to check if any of the recognizers generated had interpreted any symbols partially or completely.  These systems are less computaitonal expensive than graph matching approach but they also operate based on the output of the segmentation process. Any error in the segmentation process lead to huge error while parsing the whole sketch. Moreover, symbols cannot share strokes between them which infleunce and restrict how the user draw symbols. 
 % complexity of the graph matching problem 
\begin{figure}
	\centering
		\subfigure[An Arrow]
			{\label{fig:arrow1}\includegraphics[scale=0.75]{../../neededfiles/Figures/arrowladderdef1.eps}}
			\hfill
		\subfigure[Arrow grammar definition]
		 {\label{fig:arr2}\includegraphics[scale=0.88]{../../neededfiles/Figures/arrowladderdef2.eps}}
	\caption[Arrow definition]{}
		\label{fig:arrowladderdef1}
\end{figure}

@@

%Zernike moments were used by Hse et al. in \cite{ToADD}. 
Hse et al. \cite{HeloiseBeautification,zernike61} proposed using the magnitudes of Zernike moments as a shape property. Zernike moemnts are orthognals moments which define the distribution of point in the input space. The higher the order of the moment the finer the detail level the point distribution represents. Hse et al. \cite{HeloiseBeautification,zernike61} used a SVM classifier to train the features vector calculated from the moments. Early evluation of Zernike moemnts showed that they are invariant to both reflection and rotation of the input shape \cite{ZerMomentOrthogonal}. In our system we used Zernike moemnts as part of our feature vector. A detailed comparison between Zernike moments and other features are demonstrated in Section \ref{sec:featuresComparisions}.

@@

In all of these system only \cite{SketchRead2007} using stroke based recognitiion have been implemented to the fully creat a system that free draw sketch and understand them. This is because the emphasis on the interactive side of the sketch sytem which result in large difficultes in processing free hand sketches. Methods based on gesture and global stroke properties produce systems that are dependent on assumptions about how strokes can be groubed. These systems assume that users will finish drawing a shape before moving to the next one \cite{multistrokeEvaluation,Cali63,geometrydomain49}. This let them to temporally group strokes and then recognize them based on their strokes or apperance properties. \cite{AlvaradoDigital} argued that this assumption dose not hold in natural sketchs.

@@

\section{Swram Optimization Review}
\label{sec:ParticleSwramOptimizationReview}

%This seems to emerge of what is called 'collective intelligence' from often large simple agents.  
% start with swarm in general  .. say about flock , ant and usin in image processing 
% then speeak about PSO and  its firstly generated   (see article)
%mention that is best than ga (see thesis)

Modeling how a biological swarm has been recentlsy studied by biologist and computer scientists. Scientist in the field or artifical life need to understand how such "social anaimals" interact, achieve goals and evolove. Moreover, swarm behaviours that achieves some kind of swram intelligance is increasly intersting for engineers as it they can be applied in optimization. Looking from a higher view, agents in a swarm are cooperating to achieve some proposeful behaviour and achieve some common goal. For example, ant groups looking for food source. Large gourps of simple independent agents seems to generate this behaviour in what is called 'collective intelligence'. Each agent has a set of simple rules to follow, and accourding to the interaction of the entrie group the objective is met\cite{swarmGeneral}. %with other neighbourhood agents.

Reynolds \cite{swarmflock} created 'boid' which simulated the motion of a flock of birds. Each boid is a simple agent that navigate according to his perception and a set of basic rules.  Kube and Zhange \cite{antsimulation} introduced a simulation model for ants behvaiour and used in cooperative transportation. Ants leaves a trail of its movement, so when an ant find food it cannot cary alone, it can easily trace back to the colony and then a group of ants can track the trail and bring the food back.   Kube and Zhange \cite{antsimulation,swarmGeneral} system seems to corectly models ants behavour with a methods of task modeling.

Image processing is one of the displines that used Swarm Intelligance. Ant colony was widely used to segment objects in \cite{antSegmentation,antthreshold,SwarmMedical}. Since 1995 when Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced particle swarm alogrithm (PSO) the algorithm had been used in various image processing sytem as an optimization  algorithm \cite{SwarmImageProcessing,Swarmimageprocessing2,Colorquantization,SwarmMedical}.  PSO algorithm was used in \cite{ToADD} to classify pixel to simplify the object segmentaiton process.

ToADD
